---
title: "Peer Assessment 1"
author: "Tatiana"
date: "Sunday, May 17, 2015"
output: html_document
---

*Assignment 1 for Reproducible research course*

First step is to download the data and create the data frame for the analysis

```{r getting the data}
act_data<-read.csv("C://Users//пользователь//Downloads//activity.csv",header=TRUE,sep=",")
act_data$date<-as.POSIXct(act_data$date, format="%Y-%m-%d")
```

Now we can analyze our data and answer the following questions:

1. What is mean total number of steps taken per day?

2. What is the average daily activity pattern?

3. Are there differences in activity patterns between weekdays and weekends?

**What is mean total number of steps taken per day?**

```{r number of steps per day}
dates<-data.frame(row=1:61,date=c(unique(act_data$date)))
steps<-by(act_data$steps,act_data$date,sum,na.rm=TRUE)
stepsbydate<-data.frame(as.POSIXct(dates[,2], format="%Y-%m-%d"),c(steps))
row.names(stepsbydate)<-NULL
mean(stepsbydate[,2],na.rm=TRUE)
median(stepsbydate[,2],na.rm=TRUE)

```

Now we can plot a histogram of the total number of steps taken each day

```{r hist of steps, echo=FALSE}
hist(stepsbydate$c.steps,breaks = seq(0,22000,1000),col="grey",
main="Total number of steps taken each day (October-November 2012)",
xlab="Total number of steps")

```

**What is the average daily activity pattern?**

To establish that we need to calculate the mean of steps taken in every 5 minute interval and make a time series plot.

```{r average daily activity pattern, echo=FALSE}
interval<-data.frame(Row=1:288,period=c(unique(act_data$interval)))
steps <- by(act_data$steps, act_data$interval, mean,na.rm=TRUE)
stepsbyinterval <- data.frame(interval=interval[,2], steps=c(steps))
row.names(stepsbyinterval)<-NULL
plot(x=stepsbyinterval$interval,y=stepsbyinterval$steps,type="l",
     main = "Average Number of Steps Taken by 5-minute interval",
     xlab="Number of Interval",ylab="Step count")
```

We also need to know which interval comtains the maximum number of steps

```{r maximum number of steps}
max(stepsbyinterval$steps,na.rm=TRUE)
stepsbyinterval$interval[stepsbyinterval$steps == max(stepsbyinterval$steps)]
```

Our data frame contains some missimg values (NA), which may affect the analysis.
First we need to find how many missing values we have.

```{r missing values,echo=TRUE}
sum(is.na(act_data$steps))
```

In the new data set missing values were filled in with the mean for that day.

```{r filling in the missing values, echo=FALSE}
act_data2 <- transform(act_data, steps = ifelse(is.na(steps), ave(steps, interval, FUN = function(x) mean(x, na.rm = TRUE)), steps))
```

To compare two data sets we calculated the mean and median for the new data set and created histogram of the total number of steps taken each day

```{r new data histogram,echo=TRUE}
steps2<-by(act_data2$steps,act_data2$date,sum,na.rm=TRUE)
stepsbydate2<-data.frame(as.POSIXct(dates[,2], format="%Y-%m-%d"),c(steps2))
row.names(stepsbydate2)<-NULL
mean(stepsbydate2[,2],na.rm=TRUE)
median(stepsbydate2[,2],na.rm=TRUE)
hist(stepsbydate2$c.steps,breaks = seq(0,22000,1000),col="grey",
main="Total number of steps taken each day (October-November 2012)",
xlab="Total number of steps")
```

**Are there differences in activity patterns between weekdays and weekends?**

First we need to add a column specifing whether day was week day of weekend
```{r adding day of week,echo=TRUE}
list_of_days <- as.Date(stepsbydate2[,1])
names_of_days <- weekdays(list_of_days)
act_data2$wday <- weekdays(as.Date(act_data2$date))
d <- ifelse(act_data2$wday %in% c('суббота','воскресенье'), 
            "weekend", "weekday")
act_data2$wknd_or_not <- d
```

Now we can divide the data frame on two: first one will contain information about weekdays and the second - about weekends

```{r dividing the data,echo=TRUE}
act_data2weekday <- subset(act_data2, wknd_or_not == "weekday")
act_data2weekend <- subset(act_data2, wknd_or_not == "weekend")
```

Next we can calculate the average number of steps taken for the 5-minute intervals and make the plot
```{r making the plot,echo=FALSE}
interval2<-data.frame(Row=1:288,period=c(unique(act_data2weekday$interval)))
steps3 <- by(act_data2weekday$steps, act_data2weekday$interval, mean,na.rm=TRUE)
stepsbyinterval2 <- data.frame(interval=interval2[,2], steps=c(steps3))
row.names(stepsbyinterval2)<-NULL
interval3<-data.frame(Row=1:288,period=c(unique(act_data2weekend$interval)))
steps4 <- by(act_data2weekend$steps, act_data2weekend$interval, mean,na.rm=TRUE)
stepsbyinterval3 <- data.frame(interval=interval3[,2], steps=c(steps4))
row.names(stepsbyinterval3)<-NULL
par(mfrow =c(2,1),mar = c(2, 1, 1, 1), oma = c(1, 1, 0, 0))
plot(x=stepsbyinterval2$interval,y=stepsbyinterval2$steps,type="l",
     main = "Weekday",xlab="Interval",ylab="Steps")
plot(x=stepsbyinterval3$interval,y=stepsbyinterval3$steps,type="l",
    main = "Weekend",xlab="Interval",ylab="Steps")
```

